---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CopernicusClimate <img src="man/figures/logo.svg" align="right" height="139" copyright="cc-sa" alt="logo" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/pepijn-devries/CopernicusClimate/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pepijn-devries/CopernicusClimate/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview

[The Copernicus Climate Change Service (C3S)](https://climate.copernicus.eu/) has the mission of
providing information about the past, present and future climate, as well as tools to enable
climate change mitigation and adaptation strategies.

The C3S Climate Data Store provides open and state-of-the-art climate data to scientists. This
package allows users to download data from the data store and handle it in R.

## Installation

Install latest developmental version from R-Universe:

```{r install2, eval=FALSE}
install.packages("CopernicusClimate", repos = c('https://pepijn-devries.r-universe.dev', 'https://cloud.r-project.org'))
```

## Example

TODO

```{r example}
library(CopernicusClimate)
## basic example code
```

## More of Copernicus

More R packages for exploring other Copernicus data services:

 * [CopernicusMarine](https://github.com/pepijn-devries/CopernicusMarine)
 Dedicated to marine datasets

## Code of Conduct

Please note that the CopernicusClimate project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
